Results_setup

### 2020-04-09 doc/Results_v2.rmd
working on results and shifting code to the src file

---
title: "MLH1 Results Outline"
date: "null"
output:
  word_document:
    toc: yes
    toc_depth: '5'
  pdf_document: default
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '5'
    toc_float: yes
bibliography: refs_Main_Report.bib
---

@@

<!-- 
two round of removing outliers were done. in total 84 skel_bin potential outlier observations were removed 
-->

<!--
From this figure of the male mouse average of total sc measures, Confirm the general pattern of more SC area for the higher rec strains, compared to the low. But there are some exceptions.

the low rec musc strains which are higher are czech (1) and ast (2) -- the PWD mouse which is lower 24sep14_PWD_m1

INSERT TABLE FOR THE TOTAL SC means ect

The difference between

bin is short for binary, it reports the binarized area of the SC

skel -- takes the skeletonized midline of the bin objects
-->

@@

@@

We next fit general linear models separately for 192--`r length(fomo.male_mouse.av$mouse)` males and `r length(female_mouse.avs_for.analysis$mouse)`--144 females (M4). In the male dataset, three strains significantly affect recombination rate: musculus^PWD^ ((glm; p = 7.3710^{-10} -- `r sum_MLH1.male.specific.glm.M2$coefficients[5,][4]`; effect = ~ 5 X foci), and musculus^SKIVE^ (glm; p = 0.01; `r sum_MLH1.male.specific.glm.M2$coefficients[8,][4]` -- effect = ~7 foci), and  _molossinus^MSM^_ (glm; p=`r sum_MLH1.male.specific.glm.M2$coefficients[6,][4]`; effect ~ 2 foci). These three strains point to rapid evolution in recombination rate in spermatocytes; we subsequently refer to them collectively as "high-recombination" strains.  Analysis of the female dataset points to four strains with significant effects on recombination rate: domesticus^G^ (p = 2.510^{-6} -- `r sum_MLH1.female.spc.glm.M2$coefficients[2,][4]`), molossinus^MSM^ (p = 6.2410^{-6} -- `r sum_MLH1.female.spc.glm.M2$coefficients[5,][4]`), domesticus^LEW^ (p = 0.01 -- `r sum_MLH1.female.spc.glm.M2$coefficients[3,][4]`), and musculus^PWD^ (p=0.02  -- `r sum_MLH1.female.spc.glm.M2$coefficients[4,][4]`). Strain effect sizes in females are modest in magnitude (ranging from 1 to 4 foci) compared to those in males.

@@

```{r M2.MLH1.var, echo=FALSE, include=FALSE, eval=FALSE}
#not including this model for now / brivity
#$$mouse CO metric  ~=~ subsp * sex * strain + \varepsilon$$

MLH1var_M3 <- lm( var ~ subsp * sex * strain,  data=DF.HetC.MixedModel.HQ )

#summary(MLH1var_M3)
anova(MLH1var_M3)  
as.data.frame( round(summary(MLH1var_M3)$coefficients, 6))

MLH1cv_M3 <- lm(cV ~ subsp*sex*strain, data=DF.HetC.MixedModel.HQ)

@@

anova(MLH1cv_M3)
as.data.frame( round(summary(MLH1cv_M3)$coefficients, 6))

#Q12 var
MLH1var_M3_Q12 <- lm( var ~ subsp*sex*strain,  data=Q12_mouse_table_w.Ages )

#summary(MLH1var_M3_Q12)
anova(MLH1var_M3_Q12)  
as.data.frame( round(summary(MLH1var_M3_Q12)$coefficients, 6))

#Q12 cv
MLH1cv_M3_Q12 <- lm( cV ~ subsp*sex*strain, data=Q12_mouse_table_w.Ages )

#summary(MLH1cv_M3_Q12)
anova(MLH1cv_M3_Q12)#sex large F

as.data.frame( round(summary(MLH1cv_M3_Q12)$coefficients, 6))

```{r M3.MLH1.var, echo=FALSE}

glmMLH1Var_M3 <- glm( var ~ sex * strain,  data=DF.HetC.MixedModel.HQ )
glmMLH1Var_M3_sum <- summary(glmMLH1Var_M3)
#round(glmMLH1Var_M3_sum$coefficients[,c(1,4)], 5)

@@

Counting MLH1 foci in multiple oocytes for each female and multiple spermatocytes for each male allowed us to examine determinants of the within-mouse variance in recombination rate. To do this, we considered the same models as above, but replaced mean MLH1 foci count with within-mouse variance in MLH1 foci count as the dependent variable. Sex is the only variable that significantly affects recombination rate in both the mixed model (M1) (p < smallNumber -- `r round(LmerMLH1_VAR_results$sex.results, 5)`) and general linear model (M2) (p = 2.310^{-4} -- `r round(MLH1var_M3_Q12_sum$coefficients[,4][2], 5)`). In general, females have almost twice as much variance in MLH1 foci per cell compared to males (Figure 1). Since estimates of within-mouse variance may be more susceptible to technical error from the staining protocol, we repeated the analyses using a subset of cells with higher quality scores (**criteria? Q1, see methods for certeria**).

@@

The results are similar: sex is the strongest effect (p < smallNumber -- `r round(LmerMLH1_M1_Q12_VAR_results$sex.results, 5)`, M1; p = 2.310^{-4} -- `r  round(MLH1var_M3_Q12_sum$coefficients[2,4], 5)`, M2). When both quality-curated and full datasets are considered, strain does not significantly and consistently affect variance in MLH1 foci count per cell in either sex. These results suggest that within-mouse variance in recombination rate evolves independently of mean recombination rate.

@@

l.anova.pval <- anova(mod.L)$`Pr(>F)`[1] #0.0002701 ***
z.anova.pval <- anova(mod.Z)$`Pr(>F)`[1] #NS
```

@@

The high-recombination strains have significantly more DMC1 foci than the low-recombination strains in leptotene cells (t-test, p<**someSmallNumber** -- `r ttest.HighLow.L.pval`; one-way ANOVA, p = 0.00027 -- `r format(l.anova.pval, scientific=F)`). In contrast, the two strain groups do not differ in DMC1 foci counted in zygotene cells (later prophase)(t-test, p = 0.66 -- `r ttest.HighLow.Z.pval`; one-way-anova, p = 0.15 -- `r z.anova.pval`). **note that this interpertation depends on which stage is most relevant to CO number**

>>>>

---
title: "MLH1 Results Outline"
date: 'null'
output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '5'
    toc_float: yes
  pdf_document: default
  word_document:
    toc: yes
    toc_depth: '5'
bibliography: refs_Main_Report.bib
---

@@

@@

```{r bret.comment.var, echo=FALSE}

@@

We next fit general linear models separately for 192--`r length(fomo.male_mouse.av$mouse)` males and `r length(female_mouse.avs_for.analysis$mouse)`--144 females (M4  **male specific, M2? rename the code for these models**). In the male dataset, three strains significantly affect recombination rate: musculus^PWD^ ((glm; p = 7.3710^{-10} -- `r sum_MLH1.male.specific.glm.M2$coefficients[5,][4]`; effect = 
`r MLH1.male.specific.glm.M2$coefficients[5]` -- ~ 5 foci), and musculus^SKIVE^ (glm; p = 0.01; `r sum_MLH1.male.specific.glm.M2$coefficients[8,][4]` -- effect = `r MLH1.male.specific.glm.M2$coefficients[8]` -- ~7 foci), and  _molossinus^MSM^_ (glm; p=`r sum_MLH1.male.specific.glm.M2$coefficients[6,][4]`; effect  `r MLH1.male.specific.glm.M2$coefficients[6]` ~ 2 foci). These three strains point to rapid evolution in recombination rate in spermatocytes; we subsequently refer to them collectively as "high-recombination" strains.  Analysis of the female dataset points to four strains with significant effects on recombination rate: domesticus^G^ (p = 2.510^{-6} -- `r sum_MLH1.female.spc.glm.M2$coefficients[2,][4]`), molossinus^MSM^ (p = 6.2410^{-6} -- `r sum_MLH1.female.spc.glm.M2$coefficients[5,][4]`), domesticus^LEW^ (p = 0.01 -- `r sum_MLH1.female.spc.glm.M2$coefficients[3,][4]`), and musculus^PWD^ (p=0.02  -- `r sum_MLH1.female.spc.glm.M2$coefficients[4,][4]`). Strain effect sizes in females are modest in magnitude (ranging from 1 to 4 foci) compared to those in males.

@@

```{r M2.MLH1.var, echo=FALSE, include=FALSE}
#not including this model for now / brivity
#$$mouse CO metric  ~=~ subsp * sex * strain + \varepsilon$$

#why are these named M3? -- rename to M2

MLH1var_M2 <- lm( var ~ subsp * sex * strain,  data=DF.HetC.MixedModel.HQ )

MLH1var_M2_sum <- summary(MLH1var_M2)

#summary(MLH1var_M2)
#anova(MLH1var_M2)  
#as.data.frame( round(summary(MLH1var_M2)$coefficients, 6)) #table of p values for coefficients.

MLH1cv_M2 <- lm(cV ~ subsp*sex*strain, data=DF.HetC.MixedModel.HQ)

@@

#anova(MLH1cv_M2)
#as.data.frame( round(summary(MLH1cv_M2)$coefficients, 6))

#Q12 var
MLH1var_M2_Q12 <- lm( var ~ subsp*sex*strain,  data=Q12_mouse_table_w.Ages )

#summary(MLH1var_M3_Q12)
#anova(MLH1var_M2_Q12)  
#as.data.frame( round(summary(MLH1var_M2_Q12)$coefficients, 6))

#Q12 cv
MLH1cv_M2_Q12 <- lm( cV ~ subsp*sex*strain, data=Q12_mouse_table_w.Ages )

#summary(MLH1cv_M2_Q12)
#anova(MLH1cv_M2_Q12)#sex large F
#as.data.frame( round(summary(MLH1cv_M2_Q12)$coefficients, 6))
```

```{r M3.MLH1.var, echo=FALSE}
glmMLH1Var_M3 <- glm( var ~ sex * strain,  data=DF.HetC.MixedModel.HQ )
glmMLH1Var_M3_sum <- summary(glmMLH1Var_M3)
#round(glmMLH1Var_M3_sum$coefficients[,c(1,4)], 5)

#use the values from this model below

@@

Counting MLH1 foci in multiple oocytes for each female and multiple spermatocytes for each male allowed us to examine determinants of the within-mouse variance in recombination rate. To do this, we considered the same models as above, but replaced mean MLH1 foci count with within-mouse variance in MLH1 foci count as the dependent variable. Sex is the only variable that significantly affects recombination rate in both the mixed model (M1) (p < smallNumber -- `r round(LmerMLH1_VAR_results$sex.results, 5)`) and general linear model (M2) (p = 2.310^{-4} -- `r round(MLH1var_M3_Q12_sum$coefficients[,4][2], 5)` (old) `r round(MLH1var_M2_sum$coefficients[,4][4], 5)`  correct). **the pland full M2 model is not significant**.  M3 `r round(glmMLH1Var_M3_sum$coefficients[,4][2], 5)`.

In general, females have almost twice as much variance in MLH1 foci per cell compared to males (Figure 1). Since estimates of within-mouse variance may be more susceptible to technical error from the staining protocol, we repeated the analyses using a subset of cells with higher quality scores (quality score 1 and 2, see Materials and Methods).

@@

The results are similar: sex is the strongest effect (p < smallNumber -- `r round(LmerMLH1_M1_Q12_VAR_results$sex.results, 5)`, M1; p = 2.310^{-4} -- `r  round(MLH1var_M3_Q12_sum$coefficients[2,4], 5)`, M2 -- also M3). When both quality-curated and full datasets are considered, strain does not significantly and consistently affect variance in MLH1 foci count per cell in either sex. These results suggest that within-mouse variance in recombination rate evolves independently of mean recombination rate.

@@

l.anova.pval <- anova(mod.L)$`Pr(>F)`[1] #0.0002701 ***
z.anova.pval <- anova(mod.Z)$`Pr(>F)`[1] #NS

#anova's aren't appropriate for 2 samples... re-do the anova
#so that it's compareing more than just 2 categories -- or delete

@@

The high-recombination strains have significantly more DMC1 foci than the low-recombination strains in leptotene cells (t-test, p<**someSmallNumber** -- `r ttest.HighLow.L.pval`; one-way ANOVA, p = 0.00027 -- `r format(l.anova.pval, scientific=F)`). In contrast, the two strain groups do not differ in DMC1 foci counted in zygotene cells (later prophase)(t-test, p = 0.66 -- `r ttest.HighLow.Z.pval`). **note that this interpertation depends on which stage is most relevant to CO number**

### 2020-04-09 src/Result_Setup.Rmd
working on results and shifting code to the src file

>>>>

# MLH1 models

M1 - subsp, sex, (1 | strain)
M2 - subsp, sex, strain
M3 - sex, strain
M4 - strain (sex specific)

#MLH1 variance models

M1 - subsp, sex, (1 | strain)
M2 - subsp, sex, strain
M3 - sex, strain
M4 - strain (sex specific)

```{r M1.MLH1.var_cv, echo=FALSE, warning=FALSE, message=FALSE}
MLH1var_M1 <- lmer( var ~ subsp * sex +(1|strain), data=DF.HetC.MixedModel.HQ )
MLH1var_M1.reduce.sex <- lmer( var ~ subsp  +(1|strain), data=DF.HetC.MixedModel.HQ )
MLH1var_M1.reduce.subp <- lmer( var ~ sex  +(1|strain), data=DF.HetC.MixedModel.HQ )

#summary(MLH1var_M1)
LmerMLH1_VAR_int <- drop1(MLH1var_M1, test="Chisq")$`Pr(Chi)`[2]
LmerMLH1_VAR_sex <- anova(MLH1var_M1, MLH1var_M1.reduce.sex)$`Pr(>Chisq)`[2]
LmerMLH1_VAR_subsp <- anova(MLH1var_M1, MLH1var_M1.reduce.subp)$`Pr(>Chisq)`[2]
LmerMLH1_VAR_strain <- exactRLRT(MLH1var_M1)$p.value

LmerMLH1_VAR_results <- data.frame(sub.sp = LmerMLH1_VAR_subsp,
sex.results =  LmerMLH1_VAR_sex, interaction.pval = LmerMLH1_VAR_int,  rand.pval = LmerMLH1_VAR_strain)

```{r M2.MLH1.var, echo=FALSE, include=FALSE}
#not including this model for now / brivity
#$$mouse CO metric  ~=~ subsp * sex * strain + \varepsilon$$

MLH1var_M2 <- lm( var ~ subsp * sex * strain,  data=DF.HetC.MixedModel.HQ )
MLH1var_M2_sum <- summary(MLH1var_M2)

#summary(MLH1var_M2)
#anova(MLH1var_M2)  
#as.data.frame( round(summary(MLH1var_M2)$coefficients, 6)) #table of p values for coefficients.

MLH1cv_M2 <- lm(cV ~ subsp*sex*strain, data=DF.HetC.MixedModel.HQ)

#summary(MLH1cv_M3)
#anova(MLH1cv_M2)
#as.data.frame( round(summary(MLH1cv_M2)$coefficients, 6))

#Q12 var
MLH1var_M2_Q12 <- lm( var ~ subsp*sex*strain,  data=Q12_mouse_table_w.Ages )

#summary(MLH1var_M3_Q12)
#anova(MLH1var_M2_Q12)  
#as.data.frame( round(summary(MLH1var_M2_Q12)$coefficients, 6))

#Q12 cv
MLH1cv_M2_Q12 <- lm( cV ~ subsp*sex*strain, data=Q12_mouse_table_w.Ages )

#summary(MLH1cv_M2_Q12)
#anova(MLH1cv_M2_Q12)#sex large F
#as.data.frame( round(summary(MLH1cv_M2_Q12)$coefficients, 6))
```

```{r M3.MLH1.var, echo=FALSE}

glmMLH1Var_M3 <- glm( var ~ sex * strain,  data=DF.HetC.MixedModel.HQ )
glmMLH1Var_M3_sum <- summary(glmMLH1Var_M3)
#round(glmMLH1Var_M3_sum$coefficients[,c(1,4)], 5)

MLH1cv_M3 <- glm(cV ~ sex * strain, data=DF.HetC.MixedModel.HQ)
#summary(MLH1cv_M4)
#round(summary(MLH1cv_M3)$coefficients[,c(1,4)], 6)

### Q12 var
MLH1var_M3_Q12 <- glm( var ~ sex*strain, data=Q12_mouse_table_w.Ages )
MLH1var_M3_Q12_sum <- summary(MLH1var_M3_Q12)

#as.data.frame( round(summary(MLH1var_M3_Q12)$coefficients[,c(1,4)], 6))
#round(summary(MLH1var_M3_Q12)$coefficients[,c(1,4)], 6)

#Q12 cv
MLH1cv_M4_Q12 <- glm( cV ~ sex*strain, data=Q12_mouse_table_w.Ages )
#round(summary(MLH1cv_M4_Q12)$coefficients[,c(1,4)], 6)
```

```{r MLH1.var.M4}

#male

#female
```
